EDA

### 2024-11-29 other/EDA/EDA.qmd
filtering for clearer graphs
exploring small contracts (<=200k)

# Filter for military spending (where buyer is "National Defence")
public_services_df <- df %>%
  filter(buyer == "Public Services and Procurement Canada")

@@

```{r}
# Add "year" column to the dataframe
df$year <- format(df$award_date, "%Y")

@@

# Create a new column to categorize buyers
df$buyer_category <- ifelse(df$buyer %in% selected_buyers, df$buyer, "Other")

# Summarize data
stacked_df <- df %>%
  group_by(year, buyer_category) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

@@

@@

# Add a new column for buyer categories
df <- df %>%
  mutate(buyer_category = ifelse(buyer %in% selected_buyers, buyer, "Other"))

# Summarize total spending per year for each buyer category
grouped_df <- df %>%
  group_by(year, buyer_category) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

@@

```{r}
hist(df$amount,
     breaks = 30,  # Number of bins; adjust as needed
     main = "Distribution of Award Amounts",
     xlab = "Award Amount (CAD)",
     ylab = "Frequency",
     col = "skyblue",
     border = "black")

>>>>

# Filter for public services
public_services_df <- df %>%
  filter(buyer == "Public Services and Procurement Canada")

@@

```{r}
df_filtered <- df %>%
  filter(format(award_date, "%Y") != "2019")

# Add "year" column to the dataframe
df_filtered$year <- format(df$award_date, "%Y")

@@

# Create a new column to categorize buyers
df_filtered$buyer_category <- ifelse(df_filtered$buyer %in% selected_buyers, df_filtered$buyer, "Other")

# Summarize data
stacked_df <- df_filtered %>%
  group_by(year, buyer_category) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

@@

```{r}
df_filtered <- df %>%
  filter(format(award_date, "%Y") != "2019")

df_filtered$year <- format(df_filtered$award_date, "%Y")

@@

# Add a new column for buyer categories
df_filtered <- df_filtered %>%
  mutate(buyer_category = ifelse(buyer %in% selected_buyers, buyer, "Other"))

# Summarize total spending per year for each buyer category
grouped_df <- df_filtered %>%
  group_by(year, buyer_category) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)) %>%
  ungroup()

@@

```{r}
#View(df)

df_filtered <- df %>%
  filter(amount <= 200000)

# Optional: Check how many rows are left
num_less_than_200k <- nrow(df_filtered)
total_contracts <- nrow(df)

percent_less_than <- num_less_than_200k / total_contracts

print(percent_less_than)

# Step 3: Create bins spanning every 10,000 CAD up to 200,000
# Define the bin width
bin_width <- 10000  # 10,000 CAD

# Create a sequence of break points from 0 to 200,000
breaks <- seq(0, 200000, by = bin_width)

# To ensure that the upper limit includes 200,000, set 'right = FALSE'
# This means intervals are [0,10000), [10000,20000), ..., [190000,200000]
df_filtered <- df_filtered %>%
  mutate(amount_bin = cut(amount, breaks = breaks, right = FALSE, include.lowest = TRUE))

# Step 4: Count the number of contracts in each bin
contract_counts_df <- df_filtered %>%
  group_by(amount_bin) %>%
  summarize(Contract_Count = n()) %>%
  ungroup()

# View the binned counts
print(contract_counts_df)

contract_counts_df$amount_bin <- as.character(contract_counts_df$amount_bin)

# Create the bar plot
barplot(
  height = contract_counts_df$Contract_Count,
  names.arg = contract_counts_df$amount_bin,
  las = 2,  # Rotate x-axis labels to perpendicular
  col = "skyblue",
  border = "black",
  main = "Number of Contracts per 10,000 CAD Award Amount Range",
  xlab = "Award Amount (CAD)",
  ylab = "Number of Contracts",
  cex.names = 0.8  # Adjust label size for readability
)
```

df_filtered <- df %>%
  filter(buyer == "National Defence", amount <= 200000)

# Optional: Check how many rows are left
num_less_than_200k <- nrow(df_filtered)
total_contracts <- nrow(df)

percent_less_than <- num_less_than_200k / total_contracts

print(percent_less_than)

# Step 3: Create bins spanning every 10,000 CAD up to 200,000
# Define the bin width
bin_width <- 10000  # 10,000 CAD

# Create a sequence of break points from 0 to 200,000
breaks <- seq(0, 200000, by = bin_width)

# To ensure that the upper limit includes 200,000, set 'right = FALSE'
# This means intervals are [0,10000), [10000,20000), ..., [190000,200000]
df_filtered <- df_filtered %>%
  mutate(amount_bin = cut(amount, breaks = breaks, right = FALSE, include.lowest = TRUE))

# Step 4: Count the number of contracts in each bin
contract_counts_df <- df_filtered %>%
  group_by(amount_bin) %>%
  summarize(Contract_Count = n()) %>%
  ungroup()

# View the binned counts
print(contract_counts_df)

contract_counts_df$amount_bin <- as.character(contract_counts_df$amount_bin)

# Create the bar plot
barplot(
  height = contract_counts_df$Contract_Count,
  names.arg = contract_counts_df$amount_bin,
  las = 2,  # Rotate x-axis labels to perpendicular
  col = "skyblue",
  border = "black",
  main = "Number of Contracts per 10,000 CAD Award Amount Range",
  xlab = "Award Amount (CAD)",
  ylab = "Number of Contracts",
  cex.names = 0.8  # Adjust label size for readability
)
